\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2019-04-21 archive.tex
improved physics results

Inside its learned library we find interpretable parametric drawing
routines corresponding to the families of visual objects in its
training data, like polygons, circles, and spirals
(Figure~\ref{logoMaster}, middle) -- without supervision the agent
has learned the basic types of objects in its visual world. It
additionally learns more abstract visual relationships, like
rotational symmetry, which it models by abstracting out a new higher-order
function into its library.

@@

We first task \system with inferring 60 equations describing physical laws,
like closed forms for ballistic motion, inverse square laws,
expressions for calculating centers of mass, etc.,
drawing verbatim from AP and MCAT physics `cheat sheets'
along with equation guides from machine learning textbooks.
These relationships are easy to express
given concepts like vectors or forces.
Rather than `bake in' abstractions like these we
initialize the system with Lisp-style sequence manipulation primitives like
\code{map} and \code{fold}.
We automatically convert each task into a
regression problem by randomly generating real or vector valued input/outputs,
and encode vectors as lists of reals.

@@

\begin{figure}
  \centering
  \begin{tabular}{lll}\toprule
    &\multicolumn{1}{c}{Physical Laws}&\multicolumn{1}{c}{Recursive Functions}\\\midrule
    \normalsize \rotatebox[origin=c]{90}{\pop{Programs} \& Tasks}&
    % physical laws programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        Kinetic energy:\\
        $m$, $\vec{v}\mapsto \frac{1}{2}m|\vec{v}|^2$\\
        \code{1, [0 2 -1]$\to$2.5}\\
        \code{2, [2 1 3]$\to$14}\\
        \blueCode{$f(m$, $\vec{v}) = $($f_4$ (* $m$ ($f_0$ $\vec{v}$)))}

        %% Center-of-mass:\\
        %% $m_1$, $\vec{r}_1$, $m_2$, $\vec{r}_2\mapsto \frac{m_1}{m_1 + m_2}\vec{r}_1 + \frac{m_2}{m_1 + m_2}\vec{r}_2$\\
        %% \code{2, [3 0 1], 4, [0 6 0]}$\to$\code{[1 4 0.333]}\\
        %% \code{5, [7 1 6], 5, [9 2 2]}$\to$\code{[8 0.5 4]}        
      \end{tabular}\\\\
      \begin{tabular}{l}
        Coulomb's law:\\
        $q_1$, $\vec{r}_1$, $q_2$, $\vec{r}_2\mapsto \frac{q_1q_2}{|\vec{r}_1 - \vec{r}_2|_2^2} \times \frac{\vec{r}_1 - \vec{r}_2}{|\vec{r}_1 - \vec{r}_2|}$\\
        \code{4, [1 5 2], 2, [2 5 2]$\to$[-8 0 0]}\\
        \code{-1, [0 1 1], -1, [0 0 0]$\to$[0 0.5 0.5]}        \\
        \blueCode{$f(q_1$, $\vec{r}_1$, $q_2$, $\vec{r}_2) = $}\\
        \hspace{0.2cm}\blueCode{($f_7$ $q_1$ $q_2$ (zip $\vec{r}_1$ $\vec{r}_2$ ($\lambda$ (x y) (- x y))))}
      \end{tabular}\\\\
      \begin{tabular}{l}
        Integrating angle over time:\\
        $\theta_0$, $\omega_0$, $\alpha$, $t\mapsto \frac{1}{2}\alpha t^2 + \omega_0t + \theta_0$\\
        \code{1, 3, 0, 2$\to$7}\\
        \code{-5, 0, 10, 1$\to$0}\\
        \blueCode{$f(\theta_0$, $\omega_0$, $\alpha$, t$) = $($f_5$ $\omega_0$ $\theta_0$ $\alpha$ t)}
        \end{tabular}
      \end{tabular}
    &
    % recursive functions programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        Duplicate each list element:\\
        \code{[9 2]$\to $[9 9 2 2]}\\
        \code{[5 0 3]$\to $[5 5 0 0 3 3]}\\
        \code{[1 2 3 4]$\to $[1 1 2 2 3 3 4 4]}\\
        \blueCode{$f(l) = $($f_0$ ($\lambda$ (x a) (cons x (cons x a))) l nil)}        
      \end{tabular}\\\\

      \begin{tabular}{l}
        Lengths of list-of-lists:\\
        \code{[[2 1] [1 2 9]\, []]}$\to $\code{[2 3 0]}\\
        \code{[[5 2 9 3 7 5 2]]}$\to $\code{[7]}\\
        \code{[[] [] [9 8 9 9]]$\to$[0 0 4]}\\
        \blueCode{$f(\ell) = $($f_2$ ($\lambda$ (z) ($f_4$ z)) $\ell$)}
      \end{tabular}\\\\
      \begin{tabular}{l}
        1-index into list:\\
        \code{1, [9 2 3]$\to $9}\\
        \code{2, [1 5 8]$\to $5}\\
        \code{4, [0 2 8 1 5 6]$\to $1}\\
        \blueCode{$f($n$,$l$) = $($f_7$ l (+ 1 n))}
      \end{tabular}\\\\

      \end{tabular}
    \\\midrule
    \rotatebox[origin=c]{90}{\popp{Library}}&
    % physical laws library
    \begin{tabular}{l}
      \begin{tabular}{l}
        \greenCode{$f_0($v$)\,=\,$(fold 0 ($\lambda$ (x y) (+ (* x x) y)) v)}\\
        \hspace{\helpSize}($f_0$: \emph{L-2 norm})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_1($v$)\,=\,$(fold 0 ($\lambda$ (x y) (+ x y)) v)}\\
        \hspace{\helpSize}($f_1$: \emph{Sum vector components})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_2($v,u$)\,=\,$($f_1$ (zip v u ($\lambda$ (x y) (* x y))))}\\
        \hspace{\helpSize}($f_2$: \emph{Dot product})
      \end{tabular}\\
      %(lambda (fold $0 (map (lambda (*. $0 0.)) (car $0)) (lambda (lambda (zip $0 $1 (lambda (lambda (+. $0 $1)))))))) 
      \begin{tabular}{l}
        \greenCode{$f_3($vs$)\,=\,$(fold vs (map ($\lambda$ (x) (* x 0))}\\
        \phantom{\greenCode{$f_3($vs$)\,=\,$(fold vs (}}\greenCode{(car vs))}\\
        \phantom{\greenCode{$f_3($vs$)\,=\,$(}}\greenCode{($\lambda$ (v u) (zip v u ($\lambda$ (x y) (+ x y)))))}\\
        \hspace{\helpSize}($f_3$: \emph{Sum (list of) vectors})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_4($x$)\,=\,$(/ x (+ 1 1))}\\
        \hspace{\helpSize}($f_4$: \emph{Divide by two})
      \end{tabular}\\
      %      %#(lambda (lambda (lambda (lambda (+. $2 (*. $0 (+. (*. (#(lambda (/. $0 (+. 1. 1.))) $0) $1) $3))))))) $2 $3 $1 $0)))))
      \begin{tabular}{l}
        \greenCode{$f_5($v,x,a,t$)\,=\,$(+ x (* t (+ (* ($f_4$ t) a) v)))}\\
        \hspace{\helpSize}($f_5$: \emph{Integrate given first and second derivatives})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_6($v$)\,=\,$(power ($f_0$ v) ($f_4$ 1))}\\
        \hspace{\helpSize}($f_6$: \emph{n-dimensional vector length})
      \end{tabular}\\
      %(lambda (lambda (lambda (map (lambda (*. (/. (*. (/. $0 (#(lambda (fold $0 0. (lambda (lambda (+. (*. $1 $1) $0))))) $1)) $3) (#(lambda (power (#(lambda (fold $0 0. (lambda (lambda (+. (*. $1 $1) $0))))) $0) (#(lambda (/. $0 (+. 1. 1.))) 1.))) $1)) $2)) $0)
    %(lambda (lambda (lambda (map ($\lambda$ (x) (* (/ (* (/ x ($f_0$ v)) a) ($f_6$ v)) b)) v)
    \begin{tabular}{l}
      \greenCode{$f_7($a,b,v$)\,=\,$(map ($\lambda$ (x)}\\
      \greenCode{\phantom{$f_7($a,b}(* (/ (* (/ x ($f_0$ v)) a) ($f_6$ v)) b)) v)}\\
      \hspace{\helpSize}($f_7$: Inverse square law)
      \end{tabular}
    \end{tabular}

    &
    % recursive functions programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        \popp{\code{$f_0($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
        \phantom{\code{$f_0($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_0$ (cdr l))))}}\\
          \hspace{\helpSize}($f_0$: \emph{fold})
      \end{tabular}\\
      \begin{tabular}{l}
        \popp{\code{$f_1($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
        \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_1$ (n x))))}}\\
          \hspace{\helpSize}($f_1$: \emph{unfold})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_2($f$,$l$)\,=\,$($f_0$ nil l ($\lambda$ (x a) (cons (f x) a)))}\\
        \hspace{\helpSize}($f_2$: \emph{map})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_3($f$,$p$,$n$)\,=\,$($f_1$ p f (+ 1) n)}\\
        \hspace{\helpSize}($f_3$: \emph{count upward until predicate holds})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_4($l$)\,=\,$($f_0$ ($\lambda$ (a x) (+ 1 a)) l 0)}\\
        \hspace{\helpSize}($f_4$: \emph{length})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_5($p$)\,=\,$($f_3$ ($\lambda$ (x) x) p 0)}\\
        \hspace{\helpSize}($f_5$: \emph{count upward from 0 until predicate holds})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_6($n$)\,=\,$($f_5$ (eq? n))}\\
        \hspace{\helpSize}($f_6$: \emph{range})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_7($l$,$n$)\,=\,$(car ($f_0$ ($\lambda$ (a x) (cdr a)) ($f_6$ n) l))}\\
        \hspace{\helpSize}($f_7$: \emph{index})
      \end{tabular}

@@

  \end{tabular}
  \caption{\textbf{Left:} Learning a language for physical laws starting with recursive list routines like \code{map} and \code{fold}. Agent observes input/outputs of latent functions encoding physical laws (top left, 20 input/outputs per task). Vectors are represented as lists of real numbers. Physical constants expressed in Planck units. \textbf{Right:} Learning a language for recursive list routines starting with only recursion and primitives found in 1959 Lisp. Agent rediscovers \code{map}, \code{fold}, \code{unfold}, etc. For full tasks and learned libraries see Appendix~\ref{appendixMcCarthy}}\label{librariesToLanguages}
  \end{figure}

@@

\emph{(what we found)} Our goal was to build a model capturing
humanlike aspects of learning to be an expert in a new domain of
problems. Our model works through a wake/sleep algorithm that grows an increasingly deep
representation of a domain's conceptual structure,
while learning to quickly combine these concepts into programs
solving specific problems within the domain.
Two main findings of our work are that, first, a single
program-induction system can learn to solve large sets of problems
from many qualitatively different domains; and second, that fully
acquiring expertise in these domains hinges both upon explicit
declarative knowledge, and implicit procedural skill.  We have
deliberately placed the interplay of declarative and procedural
knowledge at the center of our model, and believe that both these
kinds of learning are crucial for building agents that, like humans,
autonomously learn to navigate a new domain of problems.

\emph{(related work)} \system builds on several generations of research in AI, program synthesis,
and cognitive science, with program induction being one of the oldest
theoretical frameworks for concept learning within artificial
intelligence~\cite{solomonoff1964formal}, and the conceptually allied
`Language of Thought Hypothesis' being almost as
old~\cite{fodor1975language}. Recent neural program synthesis systems
pair a fixed domain-specific programming language (a `DSL') to a
learned neural network that guides program
search~\cite{DBLP:journals/corr/abs-1804-01118,balog2016deepcoder,devlin2017neural}, while recent
symbolic AI research has developed frameworks for learning the DSL by
inferring reusable pieces of
code~\cite{ecc,Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
A main goal of this work is to combine and refine these techniques
with the intention of building agents that, like humans, develop
domain expertise for new classes of problems.  Work in cognitive
modeling has often exploited program-like representations for both
concept~\cite{piantadosi2011learning,lake2015human,GoodmanEtAl2015-Chapter} and intuitive theory
learning~\cite{logical,Ullman2012},
and we view these models of concept learning as
a kind of program induction,
and see theory learning as similar to inducing a DSL or library.

@@

\emph{(limitations - and how to go beyond them)} Our model's domain knowledge grows over a series of wake/sleep cycles,
with the solutions to easier problems bootstrapping
solving of harder tasks.
This dynamic is related to, but distinct from, curriculum learning. In curriculum learning approaches an agent solves a stream of pedagogically selected, increasingly difficult tasks; here we consider the case with the agent lacks a `teacher' and instead self-paces its way through an implicit curriculum.
But humans select their tasks in much richer ways,
and can even generate their own tasks to solve,
either as stepping stones toward hard, unsolved problems or motivated by considerations like curiosity and aesthetics.
Building agents that generate their own problems in these humanlike ways is a necessary next step
if we want machines that push against boundary of human domain knowledge or which even discover whole new domains of problems,
both of which go far beyond the relatively small-scale demonstrations in this work.

\emph{(I feel the lifelong learning setting is less important of a point to make - we might remove this)} Over a lifetime humans develop expertise in many domains --
unlike our model,
which learned to solve many classes of problems,
but only class one at a time.
Scaling approaches like ours to this multidomain, lifelong setting
could be enabled by a hierarchical Bayesian model of DSLs for many domains,
capturing  shared conceptual structure between similar domains  (e.g., Parcheesi and chess)
and separating knowledge between dissimilar domains (e.g., Parcheesi and cheese making).

>>>>

Inside its learned library we find interpretable parametric drawing
routines corresponding to the families of visual objects in its
training data, like polygons, circles, and spirals
(Figure~\ref{logoMaster}, middle) -- without supervision the agent
has learned the basic types of objects in its visual world. It
additionally learns more abstract visual relationships, like
radial symmetry, which it models by abstracting out a new higher-order
function into its library.

@@

We first task \system with inferring 60 equations describing physical laws,
like closed forms for ballistic motion, inverse square laws,
expressions for calculating centers of mass, etc.,
which we took from AP and MCAT physics `cheat sheets'
along with equation guides from machine learning textbooks.
These relationships are easy to express
given concepts like vectors or forces.
Rather than `bake in' abstractions like these we
initialize the system with Lisp-style sequence manipulation primitives like
\code{map} and \code{fold}.
We automatically convert each task into a
regression problem by randomly generating real or vector valued input/outputs,
and encode vectors as lists of reals.

@@

\begin{figure}
  \centering
  \setlength{\tabcolsep}{2pt} % default is 6pt
  \begin{tabular}{lll}\toprule
    &\multicolumn{1}{c}{Physical Laws}&\multicolumn{1}{c}{Recursive Functions}\\\midrule
    \normalsize \rotatebox[origin=c]{90}{\pop{Programs} \& Tasks}&
    % physical laws programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        Kinetic energy:\\
        $m$, $\vec{v}\mapsto \frac{1}{2}m|\vec{v}|^2$\\
        \code{1, [0 2 -1]$\to$2.5}\\
        \code{2, [2 1 3]$\to$14}\\
        %#(lambda (lambda (*. (/. $0 (+. $1 $1))))) 1. $1 (#(lambda (fold $0 0. (lambda (lambda (+. $0 (*. $1 $1)))))) $0))))
        \blueCode{$f(m$, $\vec{v}) = $($f_5$ 1 $m$ ($f_0$ $\vec{v}$))}

        %% Center-of-mass:\\
        %% $m_1$, $\vec{r}_1$, $m_2$, $\vec{r}_2\mapsto \frac{m_1}{m_1 + m_2}\vec{r}_1 + \frac{m_2}{m_1 + m_2}\vec{r}_2$\\
        %% \code{2, [3 0 1], 4, [0 6 0]}$\to$\code{[1 4 0.333]}\\
        %% \code{5, [7 1 6], 5, [9 2 2]}$\to$\code{[8 0.5 4]}        
      \end{tabular}\\\\
      \begin{tabular}{l}
        Coulomb's law:\\
        $q_1$, $\vec{r}_1$, $q_2$, $\vec{r}_2\mapsto \frac{q_1q_2}{|\vec{r}_1 - \vec{r}_2|_2^2} \times \frac{\vec{r}_1 - \vec{r}_2}{|\vec{r}_1 - \vec{r}_2|}$\\
        \code{4, [1 5 2], 2, [2 5 2]$\to$[-8 0 0]}\\
        \code{-1, [0 1 1], -1, [0 0 0]$\to$[0 0.5 0.5]}        \\
        \blueCode{$f(q_1$, $\vec{r}_1$, $q_2$, $\vec{r}_2) = $}\\
        \hspace{0.2cm}\blueCode{($f_{10}$ $q_1$ $q_2$ (zip $\vec{r}_1$ $\vec{r}_2$ ($\lambda$ (x y) (- x y))))}
      \end{tabular}\\\\
      \begin{tabular}{l}
        %% Integrating angle over time:\\
        %% $\theta_0$, $\omega_0$, $\alpha$, $t\mapsto \frac{1}{2}\alpha t^2 + \omega_0t + \theta_0$\\
        %% \code{1, 3, 0, 2$\to$7}\\
        %% \code{-5, 0, 10, 1$\to$0}\\
        %% \blueCode{$f(\theta_0$, $\omega_0$, $\alpha$, t$) = $($f_5$ $\omega_0$ $\theta_0$ $\alpha$ t)}
        Newton's Second Law:\\
        $m$, $\{\vec{F}_i \}\mapsto \frac{1}{m}\sum_i\vec{F}_i = a$\\
        \code{2, [[2 4 1]]$\to$[1 2 0.5]}\\
        \code{3, [[0 0 2] [3 0 1]]$\to$[1 0 0.333]}\\
        \blueCode{$f(m$, forces$) = $($f_4$ ($f_7$ m) ($f_3$ forces))}
        \end{tabular}
      \end{tabular}
    &
    % recursive functions programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        Duplicate each list element:\\
        \code{[9 2]$\to $[9 9 2 2]}\\
        \code{[5 0 3]$\to $[5 5 0 0 3 3]}\\
        \code{[1 2 3 4]$\to $[1 1 2 2 3 3 4 4]}\\
        \blueCode{$f(l) = $($f_0$ ($\lambda$ (x a) (cons x (cons x a))) l nil)}        
      \end{tabular}\\\\

      \begin{tabular}{l}
        Lengths of list-of-lists:\\
        \code{[[2 1] [1 2 9]\, []]}$\to $\code{[2 3 0]}\\
        \code{[[5 2 9 3 7 5 2]]}$\to $\code{[7]}\\
        \code{[[] [] [9 8 9 9]]$\to$[0 0 4]}\\
        \blueCode{$f(\ell) = $($f_2$ ($\lambda$ (z) ($f_4$ z)) $\ell$)}
      \end{tabular}\\\\
      \begin{tabular}{l}
        1-index into list:\\
        \code{1, [9 2 3]$\to $9}\\
        \code{2, [1 5 8]$\to $5}\\
        \code{4, [0 2 8 1 5 6]$\to $1}\\
        \blueCode{$f($n$,$l$) = $($f_8$ l (+ 1 n))}
      \end{tabular}\\\\

      \end{tabular}
    \\\midrule
    \rotatebox[origin=c]{90}{\popp{Library}}&
    % physical laws library
    \begin{tabular}{l}
      \begin{tabular}{l}
        \greenCode{$f_0($v$)\,=\,$(fold 0 ($\lambda$ (x y) (+ (* x x) y)) v)}\\
        \hspace{\helpSize}($f_0$: \emph{L-2 norm}) % still here
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_1($v$)\,=\,$(fold 0 ($\lambda$ (x y) (+ x y)) v)}\\
        \hspace{\helpSize}($f_1$: \emph{Sum vector components}) % still here
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_2($v,u$)\,=\,$($f_1$ (zip v u ($\lambda$ (x y) (* x y))))}\\
        \hspace{\helpSize}($f_2$: \emph{Dot product}) % still here
      \end{tabular}\\ 
      \begin{tabular}{l}
        \greenCode{$f_3($vs$)\,=\,$(fold vs (map ($\lambda$ (x) (* x 0))}\\
        \phantom{\greenCode{$f_3($vs$)\,=\,$(fold vs (}}\greenCode{(car vs))}\\
        \phantom{\greenCode{$f_3($vs$)\,=\,$(}}\greenCode{($\lambda$ (v u) (zip v u ($\lambda$ (x y) (+ x y)))))}\\
        \hspace{\helpSize}($f_3$: \emph{Sum (list of) vectors}) % still here
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_4($x,v$)\,=\,$(map ($\lambda$ (y) (* x y)) v)}\\
        \hspace{\helpSize}($f_4$: \emph{Multiply vector by scalar})
        \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_5($x,y,z$)\,=\,$(* (/ y (+ x x)) z)}
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_6($x,y$)\,=\,$(* ($f_5$ 1 x y) x)}\hspace{1cm}\greenCode{$f_7($x$)\,=\,$(/ 1 x)}
      \end{tabular}\\
      %      %#(lambda (lambda (lambda (lambda (+. $2 (*. $0 (+. (*. (#(lambda (/. $0 (+. 1. 1.))) $0) $1) $3))))))) $2 $3 $1 $0)))))
      %% \begin{tabular}{l}
      %%   \greenCode{$f_5($v,x,a,t$)\,=\,$(+ x (* t (+ (* ($f_4$ t) a) v)))}\\
      %%   \hspace{\helpSize}($f_5$: \emph{Integrate given first and second derivatives})
      %% \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_8($v$)\,=\,$(power ($f_0$ v) ($f_6$ 1 1))}\\
        \hspace{\helpSize}($f_8$: \emph{vector length})
      \end{tabular}\\
      %(lambda (lambda (lambda (map (lambda (*. (/. (*. (/. $0 (#(lambda (fold $0 0. (lambda (lambda (+. (*. $1 $1) $0))))) $1)) $3) (#(lambda (power (#(lambda (fold $0 0. (lambda (lambda (+. (*. $1 $1) $0))))) $0) (#(lambda (/. $0 (+. 1. 1.))) 1.))) $1)) $2)) $0)
      %(lambda (lambda (lambda (map ($\lambda$ (x) (* (/ (* (/ x ($f_0$ v)) a) ($f_6$ v)) b)) v)
      \begin{tabular}{l}
        \greenCode{$f_9($x,y,z$)\,=\,$(* (/ x ($f_0$ z)) y)}\\
        \hspace{\helpSize}($f_9$: \emph{scalar form of inverse square law})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_{10}($x,y,v$)\,=\,$($f_4$ ($f_9$ y (/ x ($f_7$ v)) v) v)}\\
        \hspace{\helpSize}($f_{10}$: \emph{vector form of inverse square law})
        \end{tabular}
    \end{tabular}

    &
    % recursive functions programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        \popp{\code{$f_0($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
        \phantom{\code{$f_0($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_0$ (cdr l))))}}\\
          \hspace{\helpSize}($f_0$: \emph{fold})
      \end{tabular}\\
      \begin{tabular}{l}
        \popp{\code{$f_1($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
        \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_1$ (n x))))}}\\
          \hspace{\helpSize}($f_1$: \emph{unfold})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_2($f$,$l$)\,=\,$($f_0$ nil l ($\lambda$ (x a) (cons (f x) a)))}\\
        \hspace{\helpSize}($f_2$: \emph{map})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_3($f$,$p$,$n$)\,=\,$($f_1$ p f (+ 1) n)}\\
        \hspace{\helpSize}($f_3$: \emph{count upward until predicate holds})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_4($l$)\,=\,$($f_0$ ($\lambda$ (a x) (+ 1 a)) l 0)}\\
        \hspace{\helpSize}($f_4$: \emph{length})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_5($p$)\,=\,$($f_3$ ($\lambda$ (x) x) p 0)}\\
        \hspace{\helpSize}($f_5$: \emph{count upward from 0 until predicate holds})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_6($n$)\,=\,$($f_5$ (eq? n))}\\
        \hspace{\helpSize}($f_6$: \emph{range})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_7($f$,$l$)\,=\,$($f_0$ nil l ($\lambda$ (x a) (if (f x) a (cons x a))))}\\
          \hspace{\helpSize}($f_7$: \emph{filter})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_8($l$,$n$)\,=\,$(car ($f_0$ ($\lambda$ (a x) (cdr a)) ($f_6$ n) l))}\\
        \hspace{\helpSize}($f_8$: \emph{index})
      \end{tabular}

@@

  \end{tabular}
  \setlength{\tabcolsep}{6pt} % default is 6pt
  \caption{\textbf{Left:} Learning a language for physical laws starting with recursive list routines like \code{map} and \code{fold}. Agent observes input/outputs of latent functions encoding physical laws (top left, 20 input/outputs per task). Vectors are represented as lists of real numbers. Physical constants expressed in Planck units. \textbf{Right:} Learning a language for recursive list routines starting with only recursion and primitives found in 1959 Lisp. Agent rediscovers \code{map}, \code{fold}, \code{unfold}, etc. For full tasks and learned libraries see Appendix~\ref{appendixMcCarthy}}\label{librariesToLanguages}
  \end{figure}

@@

Our goal was to build a model capturing
humanlike aspects of learning to be an expert in a new domain of
problems. Our model works through a wake/sleep algorithm that grows an increasingly deep
representation of a domain's conceptual structure,
while learning to quickly combine these concepts into programs
solving specific problems within the domain.
Two main findings of our work are that, first, a single
program-induction system can learn to solve large sets of problems
from many qualitatively different domains; and second, that fully
acquiring expertise in these domains hinges both upon explicit
declarative knowledge, and implicit procedural skill.  We have
deliberately placed the interplay of declarative and procedural
knowledge at the center of our model, and believe that both these
kinds of learning are crucial for building agents that, like humans,
autonomously learn to navigate a new domain of problems.

\system builds on several generations of research in AI, program synthesis,
and cognitive science, with program induction being one of the oldest
theoretical frameworks for concept learning within artificial
intelligence~\cite{solomonoff1964formal}, and the conceptually allied
`Language of Thought Hypothesis' being almost as
old~\cite{fodor1975language}. Recent neural program synthesis systems
pair a fixed domain-specific programming language (a `DSL') to a
learned neural network that guides program
search~\cite{DBLP:journals/corr/abs-1804-01118,balog2016deepcoder,devlin2017neural}, while recent
symbolic AI research has developed frameworks for learning the DSL by
inferring reusable pieces of
code~\cite{ecc,Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
A main goal of this work is to combine and refine these techniques
with the intention of building agents that, like humans, develop
domain expertise for new classes of problems.  Work in cognitive
modeling has often exploited program-like representations for both
concept~\cite{piantadosi2011learning,lake2015human,GoodmanEtAl2015-Chapter} and intuitive theory
learning~\cite{logical,Ullman2012},
and we view these models of concept learning as
a kind of program induction,
and see theory learning as similar to inducing a DSL or library.

@@

Our model's domain knowledge grows over a series of wake/sleep cycles,
with the solutions to easier problems bootstrapping
solving of harder tasks.
This dynamic is related to, but distinct from, curriculum learning. In curriculum learning approaches an agent solves a stream of pedagogically selected, increasingly difficult tasks; here we consider the case with the agent lacks a `teacher' and instead self-paces its way through an implicit curriculum.
But humans select their tasks in much richer ways,
and can even generate their own tasks to solve,
either as stepping stones toward hard, unsolved problems or motivated by considerations like curiosity and aesthetics.
Building agents that generate their own problems in these humanlike ways is a necessary next step
if we want machines that push against boundary of human domain knowledge or which even discover whole new domains of problems,
both of which go far beyond the relatively small-scale demonstrations in this work.

%% Over a lifetime humans develop expertise in many domains --
%% unlike our model,
%% which learned to solve many classes of problems,
%% but only class one at a time.
%% Scaling approaches like ours to this multidomain, lifelong setting
%% could be enabled by a hierarchical Bayesian model of DSLs for many domains,
%% capturing  shared conceptual structure between similar domains  (e.g., Parcheesi and chess)
%% and separating knowledge between dissimilar domains (e.g., Parcheesi and cheese making).
